---
title: "Estimation and Hypothesis Testing"
author: "Brayan Duran Medina"
output: html_document
---

# Importing libraries
```{r}
library(readxl)
library(stats)
```

# Read data from Excel file
```{r}
data <- read_excel("GSS_1972_2014.xlsx")
```

# Print column names
```{r}
colnames(data)
```

# Correct answers to eleven Science questions
```{r}
correct_answers <- c("True", "False", "True", "False", "True",
                     "False", "True", "True", "True",
                     "Earth around sun", "One year")
```

# Extract people from year 2014
```{r}
data_2014 <- data[data$"Gss year for this respondent" == 2014, ]

data_2014
```

# Extract columns names with answers to Science questions
```{r}
science_questions <- colnames(data_2014)[c(42:52)]

science_questions
```

# Transforming data
```{r}
data_2014$"Number of children" <-
  ifelse(data_2014$"Number of children" == "Eight or more",
         8, data_2014$"Number of children")
```


# Remove people that did not answer any Science question
```{r}
data_2014 <- data_2014[rowSums(data_2014[science_questions] == "Not applicable")
                       != length(science_questions), ]

data_2014
```

# Add new column with number of correct answers
```{r}
data_2014$correct_answers <- NA


for (i in seq_len(nrow(data_2014))) {
  data_2014$correct_answers[i] <-
    sum(data_2014[i, science_questions] == correct_answers)
}

data_2014$correct_answers
```

# Histogram of Number of Correct Answers
```{r}
hist(data_2014$correct_answers,
     main = "Histogram of Correct Answers",
     xlab = "Number of Correct Answers",
     ylab = "Frequency",
     col = "blue")
```

# Transform cuantitative variables categorical variables
```{r}
cuantitative_variables <- c("Age of respondent",
                            "Number of children",
                            "Highest year of school completed")

for (var in cuantitative_variables) {
  data_2014[[var]] <- as.numeric(data_2014[[var]])
}

number_of_bins <- 5

for (var in cuantitative_variables) {
  data_2014[[var]] <- cut(data_2014[[var]],
                          breaks = number_of_bins,
                          labels = c("Very Low", "Low",
                                     "Medium", "High", "Very High"))
}
```

# ANOVA test for demographic variables
```{r}

demographic_variables <- c("Respondents sex",
                           "Age of respondent",
                           "Race of respondent",
                           "Marital status",
                           "Number of children",
                           "Highest year of school completed",
                           "Sexual orientation",
                           "Respondents income",
                           "Does r own or rent home?")

significant_variables <- c()

for (var in demographic_variables) {
  anova_result <- aov(correct_answers ~ data_2014[[var]], data = data_2014)

  print(paste("ANOVA for", var))

  print(summary(anova_result))

  par(mgp = c(3, 1, 0), cex.axis = 0.6)

  boxplot(data_2014$correct_answers ~ data_2014[[var]],
          main = paste("Boxplot of Correct Answers by", var),
          xlab = "", ylab = "Number of Correct Answers", col = "lightgreen",
          las = 2)

  if (summary(anova_result)[[1]][["Pr(>F)"]][1] < 0.05) {

    significant_variables <- c(significant_variables, var)
  }
}
```

# Make a Post Hoc test for significant variables
```{r}

for (var in significant_variables) {
  anova_result <- aov(correct_answers ~ data_2014[[var]], data = data_2014)

  posthoc_result <- TukeyHSD(anova_result)

  print(paste("Post-Hoc Test for", var))

  print(posthoc_result)

  aggregate_data <- aggregate(correct_answers ~ data_2014[[var]],
                              data = data_2014, mean)

  par(mgp = c(3, 1, 0), cex.axis = 0.6)

  barplot(aggregate_data$correct_answers,
          names.arg = aggregate_data[, 1],
          main = paste("Bar Plot of Mean Correct Answers by", var),
          xlab = "", ylab = "Mean Number of Correct Answers", col = "coral",
          las = 2)
}
```

# Summary Report on the Statistical Analysis of GSS 2014 Data

## Objective

The objective of this analysis was to assess the impact of various demographic factors on scientific knowledge as indicated by the number of correct science answers in the General Social Survey (GSS) 2014 data.

## Methodology

Analysis of Variance (ANOVA) was selected to evaluate the influence of demographic variables on scientific literacy. The variables examined include respondents’ sex, age, race, marital status, number of children, educational level, sexual orientation, income, and home ownership. This choice was based on the following reasons:

Applicability for Comparing Multiple Groups: ANOVA is particularly effective when comparing means across multiple groups, which is essential in our case with diverse demographic categories.

Suitability for Categorical Independent Variables: The demographic variables in our study are categorical, making ANOVA a suitable choice for understanding their impact on a continuous dependent variable (scientific literacy).

The methodology involves analyzing each demographic factor's influence on the respondents' number of correct science answers. Significant findings from the ANOVA were further investigated using Post Hoc tests, providing a more detailed understanding of the differences between specific groups within significant demographic categories. This approach ensures a comprehensive and statistically robust analysis, shedding light on the complex interplay between various demographic factors and scientific literacy.

## Key Findings

Educational Level (Highest year of school completed):

- F value: 58.58, Pr(>F): <2e-16.
- Explanation: This strong statistical significance indicates that educational attainment is a key factor in scientific literacy. Higher education levels likely provide more comprehensive exposure to science, leading to better understanding.

Race of Respondent (Race of respondent):

- F value: 37.04, Pr(>F): 2.38e-16.
- Explanation: The high significance suggests considerable differences in scientific knowledge across racial groups. This may reflect disparities in access to quality education and cultural or societal influences.

Income Levels (Respondents income):

- F value: 7.802, Pr(>F): 4.87e-16.
- Explanation: This significant correlation indicates that higher income may facilitate better access to educational resources, leading to improved scientific literacy.

Marital Status (Marital status):

- F value: 11.74, Pr(>F): 4.12e-11.
- Explanation: Marital status appears to influence scientific knowledge, potentially due to changes in time availability and life priorities that affect learning opportunities and educational engagement.

Age of Respondent (Age of respondent):

- F value: 12.38 Pr(>F): 7.14e-10.
- Explanation: The age of respondents influences scientific literacy to a lesser extent than other factors. This could reflect generational differences in educational exposure to science or the evolving nature of science education over time.

Respondent's Sex (Respondents sex):

- F value: 36.49, Pr(>F): 2.03e-09.
- Explanation: The significant difference in scientific literacy between males and females might be attributed to gender-specific educational experiences, societal expectations, or differences in opportunities and encouragement to pursue science education.

Number of Children (Number of children):

- F value: 10.56, Pr(>F): 2.04e-08.
- Explanation: Having more children is associated with differences in scientific knowledge, possibly due to time constraints or shifting priorities that impact continuous scientific learning.

Sexual Orientation (Sexual orientation):

- F value: 5.977, Pr(>F): 1.8e-05.
- Explanation: The variation in scientific literacy among different sexual orientations suggests the influence of social dynamics, educational inclusivity, and possibly differing experiences with science education.

Home Ownership (Does r own or rent home?):

- F value: 2.577, Pr(>F): 0.025.
- Explanation: The impact of home ownership, though less pronounced, indicates underlying socioeconomic factors. Homeownership may be a proxy for financial stability, which could influence access to educational resources.

## Conclusions

This analysis has clearly shown that demographic factors significantly shape scientific literacy. The level of education, race, and income emerge as particularly influential, indicating the importance of access to quality education and the impact of socioeconomic factors on scientific understanding. Sex, marital status, and the number of children also play notable roles, reflecting how different life experiences and responsibilities can affect scientific knowledge. Additionally, variations in scientific literacy across different sexual orientations and age groups highlight the need for inclusive and diverse educational environments. These findings emphasize the importance of developing tailored educational strategies and public awareness campaigns that consider these diverse demographic influences. By addressing these factors, we can work towards enhancing scientific literacy across all segments of society, ensuring that everyone has the opportunity to engage with and understand